Appendix B: BBS Results

### 2014-08-04 improved_state_space.tex
Small revisions; added line numbers to response

\textit{Key words}: abundance, Dail and Madsen model, density-dependence,
detectability, Gompertz-logistic, immigration, % model
population dynamics, Ricker-logistic, zero-inflated % model

@@

A simple state-space model can be described as follows.
Let $N_t$ be the abundance of a species during year $t$, for
$t=1,\hdots,T$, and let $X_t$ be
the observed data, which differs from $N_t$ due to observation error,
a random effect denoted $\zeta_t$. Temporal variation in $N_t$ is
modeled using a population growth model, $\mu(N_{t-1})$
and random process variation denoted $\eta_t$.
The growth model may be density-dependent, as in the case of the 
Ricker-logistic model (henceforth Ricker model), or it might be 
density-independent, such as when growth
is exponential. %% Exponential or geometric?
A generic model can be written: 
\begin{linenomath*}
\begin{gather}
  \label{eq:ss1}
  \begin{align}
    N_1 &= x_0 \nonumber \\ %X_1
N_t &= \mu(N_{t-1}) + \eta_{t-1} \quad \text{for} \; 
t=2,\hdots,T  \\
X_t &= N_t + \zeta_t \qquad \qquad \;\, \text{for} \;
t=1,\hdots,T. \nonumber 
  \end{align}
\end{gather}
\end{linenomath*}
where $x_0$ is the model for initial abundance.  In classical
state-space models, the two sets of random effects
are assumed to be i.i.d. Gaussian variables: 
$\eta_t \sim \mathrm{N}(0, \sigma)$ and
$\zeta_t \sim \mathrm{N}(0, \tau)$. 
The process variation associated with $N_1$ is often ignored, or
the population is assumed to be at equilibrium such that the $N_1$ can
be regarded as an outcome of the equilibrium distribution.

@@

Other limitations of state-space models, as commonly applied in ecology,
include the failure to admit spatial variation in abundance and
the fact that some parameters are not identifiable
\citep{polansky_etal:2009}. 
Ignoring spatial variation reduces the scope of the
inferences, reduces the information about the parameters of interest,
and can lead to underestimation of variances \citep{dennis_etal:2010}.
Even more problematic is the issue that without some form of replication in the first year, 
abundance cannot be estimated without making strong assumptions about 
the initial distribution. 
For example, practitioners either assume that there is no
observation error in the first year ($x_0 = X_1$), or they assume the population is
at equilibrium, which may defeat the purpose of many studies of
population dynamics.

@@

\subsection*{Environmental and demographic stochasticity}
\begin{comment}
Even though the DM model is conceptually simple, 
it includes many random effects (the $N$'s)
making it challenging to compute the likelihood or implement MCMC
algorithms. Nonetheless, in some cases additional random effects
may be of interest. For example, in conventional state-space models,
environmental stochasticity is often modeled as $r_t \sim
\mathrm{Norm}(\mu_r, \sigma_r)$.
\end{comment}
The original DM models allows for demographic stochasticity --- variation in population growth 
rate due to the randomness of birth and death processes --- 
via the binomial and Poisson distributions for apparent survival and recruitment.  Our population
growth extensions use a single Poisson distribution for abundance. 
% NOTE: I would save this for the discussion
Conceptually, this is an appropriate method of modeling demographic stochasticity 
 for semelparous organisms with one generation per sampling interval (year)
 \citep{bonsall_hastings:2004},
and the DM models appropriate for modeling the demographic stochasticity of iteroparous organisms,
if they are underdispersed compared to the Poisson (but see Discussion).
% but may actually perform better for iteroparous organisms as well (see Discussion).  
 Dynamics of semelparous organisms with more than one generation per year 
 (many insects) may be overdispersed compared to the Poisson.  If the number of generations per
 year is known, then this can be accounted for in our models by adding extra $N$'s between each
 count (see Supplement 1).

@@

Bayesian inference is an alternative to classical inference with
several appealing features. First, it allows direct probability
statements to be made about a hypothesis given data
\citep{link_barker:2010}.  Second, Bayesian methods offer
straight-forward approaches for combining data from multiple sources
or use existing estimates of parameters as prior distributions. Third,
one common usage of state-space models is for predicting future
population size, and this task is easily accomplished using Bayesian
methods.

@@

\subsection*{Simulation Study to Assess Estimator Accuracy}

@@

\subsection*{Robustness Simulation Study}

We tested our models against violations of the Poisson process assumption for
annual dynamics.  We ran three alternative scenarios: 1) a geometric-recruitment model with
$\omega = \gamma = 0.5$; 2) a Leslie matrix model with six age classes; and 3) 
a Gompertz model with two non-overlapping generations per year (sampling occasion).   
%% Couldn't we expand our model to deal with these "mis-specifications"? Like Elise and I did in a couple of recent papers?
The first two scenarios are underdispersed compared to the Poisson and the third is overdispersed.
In the first two cases, we estimated parameters with both the exponential and geometric-recruitment
models using the \textbf{R} package \texttt{unmarked}.  We compared the bias,
root mean squared error, and coverage of the model estimates as well as their Akaike's Information
Criterion (AIC) values.
For the third scenario, we estimated parameters with the Gompertz model with 
one generation per year using both maximum likelihood and MCMC approaches, as well
as allowing for two generations per year with MCMC.  We compared the bias, root mean
squared error, and coverage of all three model estimates.  Details of the robustness simulation
study can be found in Appendix A.

\subsection*{Analysis of Breeding Bird Survey Data}

@@

\subsection*{Robustness Simulation Study}
In simulations of the geometric-recruitment model estimated with the exponential model,
estimates of $\Lambda$ were biased high 
and $p$ biased low (Table A5). 
However, estimates of $r$ had low bias and RMSE, with nominal coverage. 
The estimates from the geometric-recruitment model for $\Lambda$ and $p$ had low bias, but the estimates
for $\gamma$ and $\omega$ were somewhat biased and imprecise (Table A5 and Fig. A3).
The geometric- recruitment model had lower
AIC 71.3\% of the time; when the geometric-recruitment model was favored, the
mean $\Delta$AIC was 6.0 and when the exponential model was favored
the mean was 1.4.  
For the Leslie matrix simulations, bias in estimates of $\Lambda$ and $p$ from both the exponential and geometric-recruitment models
were high, but the exponential biases were higher (Table A6 and Fig. A3).  
The geometric-recruitment model was favored in 100\% of the simulations, 
with a mean $\Delta$AIC of 89.0.

@@

\subsection*{Analysis of Breeding Bird Survey Data}

@@

One strategy for adhering to the original formulation of the
model that included survival and recruitment parameters would be to
use either prior information or additional data \citep{zipkin_etal:2014}. For example, in a 
Bayesian analysis existing estimates of demographic parameters
could be used as informative priors, which might make it possible to
seperate the effects of movement versus births and deaths.
Alternatively, ecologists could combine count data with data
from, for example,
capture-recapture data. Models that are fitted to count data and
demographic data are often referred to as integrated population models
\citep[IPM;][]{besbeas_etal:2002,schaub_etal:2007,chandler_clark:2014}.
We believe that a DM modeling framework could work well in this context,
and that the DM model could be viewed as a specific case of the
IPM proposed by \citet{buckland_etal:2004}.
However, we caution that when count data are affected by
movement and vital rates, it would be unwise to ignore the movement
processes.

@@

Our models generally did not estimate parameters well in simulations when data were underdispersed
or overdispersed compared to the model assumptions.  However, we present models for
both Poisson distributed and overdispersed data, and the original DM model and the 
stage-structured model of \citet{zipkin_etal:2014} are available for underdispersed data.  AIC model
selection did a good job of picking the most appropriate distribution.  Justifiable 
density-dependent models
for underdispersed data remain to be developed.

>>>>

\textit{Key words}: abundance, Dail and Madsen model, density-dependence,
detectability, Gompertz model, immigration, %-logistic
population dynamics, Ricker model, zero-inflated %-logistic

@@

A simple state-space model can be described as follows.
Let $N_t$ be the abundance of a species during year $t$, for
$t=1,\hdots,T$, and let $X_t$ be
the observed data, which differs from $N_t$ due to observation error,
a random effect denoted $\zeta_t$. Temporal variation in $N_t$ is
modeled using a population growth model, $\mu(N_{t-1})$
and random process variation denoted $\eta_t$.
The growth model may be density-dependent, as in the case of the 
Ricker-logistic model (henceforth Ricker model), or it might be 
density-independent, such as when growth
is exponential. %% Exponential or geometric?
A generic model can be written: 
\begin{linenomath*}
\begin{gather}
  \label{eq:ss1}
  \begin{align}
    N_1 &= x_0 \nonumber \\ %X_1
N_t &= \mu(N_{t-1}) + \eta_{t-1} \quad \text{for} \; 
t=2,\hdots,T  \\
X_t &= N_t + \zeta_t \qquad \qquad \;\, \text{for} \;
t=1,\hdots,T. \nonumber 
  \end{align}
\end{gather}
\end{linenomath*}
where $x_0$ is the distribution for initial abundance.  In classical
state-space models, the two sets of random effects
are assumed to be i.i.d. Gaussian variables: 
$\eta_t \sim \mathrm{N}(0, \sigma)$ and
$\zeta_t \sim \mathrm{N}(0, \tau)$. 
The process variation associated with $N_1$ is often ignored ($x_0 = X_1$), or
the population is assumed to be at stochastic equilibrium such that the $x_0$ can
be regarded as an outcome of the equilibrium distribution \citep{devalpine_hastings:2002}.

@@

Other limitations of state-space models, as commonly applied in ecology,
include the failure to admit spatial variation in abundance and
the fact that some parameters are not identifiable
\citep{polansky_etal:2009}. 
Ignoring spatial variation reduces the scope of the
inferences, reduces the information about the parameters of interest,
and can lead to underestimation of variances \citep{dennis_etal:2010}.
Even more problematic is the issue that without some form of replication in the first year, 
abundance cannot be estimated without making strong assumptions about 
the initial distribution. 
For example, practitioners either assume that there is no
observation error in the first year, or they assume the population is
at equilibrium, which may defeat the purpose of many studies of
population dynamics.

@@

\subsection*{Environmental and demographic stochasticity}
\label{subsec:stoch}
The original DM models allows for demographic stochasticity --- variation in population growth 
rate due to the randomness of birth and death processes --- 
via the binomial and Poisson distributions for apparent survival and recruitment.  Our population
growth extensions use a single Poisson distribution for abundance. 
% NOTE: I would save this for the discussion
Conceptually, this is an appropriate method of modeling demographic stochasticity 
 for semelparous organisms with one generation per sampling interval (year)
 \citep{bonsall_hastings:2004},
and the DM models appropriate for modeling the demographic stochasticity of iteroparous organisms,
if they are underdispersed compared to the Poisson (but see Discussion).
% but may actually perform better for iteroparous organisms as well (see Discussion).  
 Dynamics of semelparous organisms with more than one generation per year 
 (many insects) may be overdispersed compared to the Poisson.  If the number of generations per
 year is known, then this can be accounted for in our models by adding extra $N$'s between each
 count (see Supplement 1).

@@

Bayesian inference is an alternative to classical inference with
several appealing features. First, it allows direct probability
statements to be made about a hypothesis given data
\citep{link_barker:2010}.  Second, Bayesian methods offer
straight-forward approaches for combining data from multiple sources
or using existing estimates of parameters as prior distributions. Third,
a common usage of state-space models is to predict future
population size, and this is easily accomplished using Bayesian
methods.

@@

\subsection*{Simulation study to assess estimator accuracy}

@@

\subsection*{Robustness simulation study}

Most of our DM model extensions assume Poisson process error for annual dynamics.
Although we provide a variation for data overdispersed compared to the Poisson (see \textit{Environmental and demographic stochasticity}), 
and the original DM models assume underdispersed data,  
we tested our models against violations of the Poisson assumption.  
We ran three alternative scenarios: 1) a geometric-recruitment model with
$\omega = \gamma = 0.5$; 2) a Leslie matrix model with six age classes; and 3) 
a Gompertz model with two non-overlapping generations per year (sampling occasion).   
The first two scenarios are underdispersed compared to the Poisson and the third is overdispersed.
In the first two cases, we estimated parameters with both the exponential and geometric-recruitment
models using the \textbf{R} package \texttt{unmarked}.  We compared the bias,
root mean squared error, and coverage of the model estimates as well as their Akaike's Information
Criterion (AIC) values.
For the third scenario, we estimated parameters with the Gompertz model with 
one generation per year using both maximum likelihood and MCMC approaches, as well
as allowing for two generations per year with MCMC.  We compared the bias, root mean
squared error, and coverage of all three model estimates.  Details of the robustness simulation
study can be found in Appendix A.

\subsection*{Analysis of Breeding Bird Survey data}

@@

\subsection*{Robustness simulation study}
In simulations of the geometric-recruitment model estimated with the exponential model,
estimates of $\Lambda$ were biased high 
and $p$ biased low (Table A5). 
However, estimates of $r$ had low bias and RMSE, with nominal coverage. 
The estimates from the geometric-recruitment model for $\Lambda$ and $p$ had low bias, but the estimates
for $\gamma$ and $\omega$ were somewhat biased and imprecise (Table A5 and Fig. A3).
The geometric- recruitment model had lower
AIC 71.3\% of the time; when the geometric-recruitment model was favored, the
mean $\Delta$AIC was 6.0 and when the exponential model was favored
the mean was 1.4.  
For the Leslie matrix simulations, bias in estimates of $\Lambda$ and $p$ from both the exponential and geometric-recruitment models
were high, but the exponential biases were higher (Table A6 and Fig. A3).  
The geometric-recruitment model was favored in 100\% of the simulations, 
with a mean $\Delta$AIC of 89.0.

@@

\subsection*{Analysis of Breeding Bird Survey data}

@@

One strategy for adhering to the original formulation of the
model that included survival and recruitment parameters would be to
use either prior information or additional data \citep{zipkin_etal:2014}. For example, in a 
Bayesian analysis existing estimates of demographic parameters
could be used as informative priors, which might make it possible to
seperate the effects of movement versus births and deaths.
Alternatively, ecologists could combine count data with demographic data,
%from, for example,
%capture-recapture data. Models that are fitted to count data and
%demographic data are 
often referred to as an integrated population model
\citep[IPM;][]{besbeas_etal:2002,schaub_etal:2007,chandler_clark:2014}.
We believe that a DM modeling framework could work well in this context,
and that the DM model could be viewed as a specific case of the
IPM proposed by \citet{buckland_etal:2004}.
However, we caution that when count data are affected by
movement and vital rates, it would be unwise to ignore the movement
processes.

@@

Our models generally did not estimate parameters well in simulations when data were underdispersed
or overdispersed compared to the model assumptions.  However, we present models for
both Poisson distributed and overdispersed data, and the original DM model and the 
stage-structured model of \citet{zipkin_etal:2014} are available for underdispersed data.  AIC model
selection did a good job of picking the most appropriate distribution.  Appropriate 
density-dependent models for underdispersed data remain to be developed.

### 2014-08-04 supp/bbs_estimates_appendix.tex
Small revisions; added line numbers to response

\begin{document}
\title{Appendix B: BBS Results}
\maketitle
\renewcommand*\thetable{B\arabic{table}}
\renewcommand*\thefigure{B\arabic{figure}}
\begin{table}[htb]
  \centering
  \small
\caption{Model selection table for ovenbirds in Maryland and Virginia,
    1966-2010.  We present model name and number, number of 
parameters (Par.), and difference in Akaike's
information criterion between each model and the top model of
that set ($\Delta$AIC).  The first section compares
models for initial abundance, the second for detection
probability, and the third for dynamics.}
  \begin{tabular}[h]{lcc}
\hline
Model	&Par.	&$\Delta$AIC	\\
\hline
A. Initial Abundance && \\
A.1. NB[$\Lambda$(.)$\alpha$(.)]Exponential[$r$(.)]$p$(.)	&4	&0\\
A.2. P[$\Lambda$(.)]Exponential[$r$(.)]$p$(.)	&3	&1262.7\\
A.3. ZIP[$\Lambda$(.)$\psi$(.)]Exponential[$r$(.)]$p$(.)	&4 &1264.7\\
\hline
B. Detection Probability && \\
B.1. NB[$\Lambda$(.)$\alpha$(.)]Exponential[$r$(.)]$p$(wind+1st)	&8 &0	\\
B.2. NB[$\Lambda$(.)$\alpha$(.)]Exponential[$r$(.)]$p$(wind) &7 &0.9\\
B.3. NB[$\Lambda$(.)$\alpha$(.)]Exponential[$r$(.)]$p$(1st)	&5	&5.0
\\B.4. NB[$\Lambda$(.)$\alpha$(.)]Exponential[$r$(.)]$p$(.)	&4	&6.4\\
\hline
C. Dynamics && \\
C.1. NB[$\Lambda$(.)$\alpha$(.)]Ricker-logistic+Immigration[$r$(.)$K$(.)$\iota$(.)]$p$(wind+1st) &10	&0	\\
C.2. NB[$\Lambda$(.)$\alpha$(.)]Gompertz-logistic+Immigration[$r$(.)$K$(.)$\iota$(.)]$p$(wind+1st) &10	&8.4 \\
C.3. NB[$\Lambda$(.)$\alpha$(.)]Exponential+Immigration[$r$(.)$\iota$(.)]$p$(wind+1st) &9	&36.5\\
C.4. NB[$\Lambda$(.)$\alpha$(.)]Geometric-recruitment+Immigration[$\gamma$(.)$\omega$(.)$\iota$(.)]$p$(wind+1st) &10	&38.6\\
C.5. NB[$\Lambda$(.)$\alpha$(.)]Gompertz-logistic[$r$(.)$K$(.)]$p$(wind+1st) &9	&192.8\\
C.6. NB[$\Lambda$(.)$\alpha$(.)]Ricker-logistic[$r$(.)$K$(.)]$p$(wind+1st) &9	&195.1\\
C.7. NB[$\Lambda$(.)$\alpha$(.)]Exponential[$r$(.)]$p$(wind+1st)	&8 &271.3\\
C.8. NB[$\Lambda$(.)$\alpha$(.)]Geometric-recruitment[$\gamma$(.)$\omega$(.)]$p$(wind+1st) &9	&273.7\\
C.9. NB[$\Lambda$(.)$\alpha$(.)]Constant-recruitment[$\gamma$(.)$\omega$(.)]$p$(wind+1st) &9	&1856.7\\
\hline
\end{tabular}
\end{table}
\clearpage

%[Andy and Richard, which do you prefer for parameter estimates?  
%Table or figure?  Either one may need a bit of work still.]
%% latex table generated in R 3.0.0 by xtable 1.7-1 package
%% Wed Sep 18 14:26:41 2013
%\begin{table}[ht]
%\centering
%  \caption{Parameter estimates for ovenbirds in Maryland and Virginia from BBS data, 1966-2010.  
%MLE estimates come from the NB[$\Lambda$(.)$\alpha$(.)]Ricker-logistic+Immigration[$r$(.)$K$(.)$\iota$(.)]$p$(wind+1st) 
%model run in the \textbf{R} package \texttt{unmarked}; base estimates from the same model run in \textbf{JAGS};
%obs estimates from the base model with random observer effects added; and ES estimates from the base
%model with regional environmental stochasticity added.  Detectability parameters (intercept: $p_{0}$, 
%effect of wind speed 1: $p_{1}$, effect of wind speed 2: $p_{2}$, effect of wind speed 3 or higher: $p_{3}$, 
%effect of first run of a route by an observer: $p_{1st}$, and random observer effect: $\sigma_{p}$) are on the 
%logit scale.}
%\begin{tabular}{rllllll}
%  \hline
% & Parameter & MLE & base & obs & ES & ES and obs \\ 
%  \hline
%1 & $\Lambda$ & 33.4 $\pm$ 4.7 & 31.1 $\pm$ 5.2 & 42.5 $\pm$ 7.3 & 33.2 $\pm$ 5.8 & 43.8 $\pm$ 7.9 \\ 
%  2 & $\alpha$ & 0.654 $\pm$ 0.087 & 0.624 $\pm$ 0.108 & 0.730 $\pm$ 0.119 & 0.649 $\pm$ 0.110 & 0.753 $\pm$ 0.124 \\ 
%  3 & $r$ & 0.026 $\pm$ 0.006 & 0.011 $\pm$ 0.005 & 0.018 $\pm$ 0.006 & 0.020 $\pm$ 0.010 & 0.023 $\pm$ 0.010 \\ 
%  4 & $K$ & 56.9 $\pm$ 9.4 & 55.8 $\pm$ 18.2 & 101.8 $\pm$ 21.1 & 112.0 $\pm$ 58.8 & 141.5 $\pm$ 69.5 \\ 
%  5 & $\iota$ & 0.283 $\pm$ 0.043 & 0.333 $\pm$ 0.043 & 0.306 $\pm$ 0.051 & 0.329 $\pm$ 0.046 & 0.303 $\pm$ 0.052 \\ 
%  6 & $p_{0}$ & -1.492 $\pm$ 0.100 & -1.476 $\pm$ 0.097 & -1.994 $\pm$ 0.116 & -1.571 $\pm$ 0.099 & -2.056 $\pm$ 0.124 \\ 
%  7 & $p_{1}$ & -0.012 $\pm$ 0.021 & -0.011 $\pm$ 0.021 & -0.014 $\pm$ 0.021 & -0.013 $\pm$ 0.021 & -0.015 $\pm$ 0.021 \\ 
%  8 & $p_{2}$ & -0.075 $\pm$ 0.027 & -0.071 $\pm$ 0.027 & -0.056 $\pm$ 0.026 & -0.069 $\pm$ 0.027 & -0.055 $\pm$ 0.026 \\ 
%  9 & $p_{3}$ & -0.133 $\pm$ 0.049 & -0.135 $\pm$ 0.049 & -0.145 $\pm$ 0.048 & -0.141 $\pm$ 0.049 & -0.144 $\pm$ 0.048 \\ 
%  10 & $p_{1st}$ & -0.058 $\pm$ 0.028 & -0.047 $\pm$ 0.027 & -0.075 $\pm$ 0.029 & -0.045 $\pm$ 0.028 & -0.075 $\pm$ 0.029 \\ 
%  31 & $\sigma_{p}$ & - & - & 0.353 $\pm$ 0.034 & - & 0.349 $\pm$ 0.034 \\ 
%  42 & $\sigma_{\nu}$ & - & - & - & 0.074 $\pm$ 0.014 & 0.064 $\pm$ 0.014 \\ 
%   \hline
%\end{tabular}
%\end{table}

\clearpage
\begin{figure}
  \centering
  \includegraphics[width=6.5in]{../figs/oven_par_est}
\caption{Parameter estimates for ovenbirds in Maryland and Virginia from BBS data, 1966-2010.  
MLE estimates come from the NB[$\Lambda$(.)$\alpha$(.)]Ricker-logistic+Immigration[$r$(.)$K$(.)$\iota$(.)]$p$(wind+1st) 
model run in the \textbf{R} package \texttt{unmarked}; base estimates from the same model run in \textbf{JAGS};
obs estimates from the base model with random observer effects added; and ES estimates from the base
model with regional environmental stochasticity added.  Detectability parameters (intercept: $p_{0}$, 
effect of wind speed 1: $p_{1}$, effect of wind speed 2: $p_{2}$, effect of wind speed 3 or higher: $p_{3}$, 
effect of first run of a route by an observer: $p_{1st}$, and random observer SD: $\sigma_{p}$) are on the 
logit scale. Error bars show SE for MLE estimates and SD for all others.}
\label{fig:oven_par_est}
\end{figure}

>>>>

\begin{document}
\title{Appendix B: BBS Results}
\date{} %remove date
\maketitle
\renewcommand*\thetable{B\arabic{table}}
\renewcommand*\thefigure{B\arabic{figure}}
\begin{table}[htb]
  \centering
  \small
\caption{Model selection table for ovenbirds in Maryland and Virginia,
    1966-2010.  We present model name and number, number of 
parameters (Par.), and difference in Akaike's
information criterion between each model and the top model of
that set ($\Delta$AIC).  The first section compares
models for initial abundance, the second for detection
probability, and the third for dynamics.}
  \begin{tabular}[h]{lcc}
\hline
Model	&Par.	&$\Delta$AIC	\\
\hline
A. Initial Abundance && \\
A.1. NB[$\Lambda$(.)$\alpha$(.)]Exponential[$r$(.)]$p$(.)	&4	&0\\
A.2. P[$\Lambda$(.)]Exponential[$r$(.)]$p$(.)	&3	&1262.7\\
A.3. ZIP[$\Lambda$(.)$\psi$(.)]Exponential[$r$(.)]$p$(.)	&4 &1264.7\\
\hline
B. Detection Probability && \\
B.1. NB[$\Lambda$(.)$\alpha$(.)]Exponential[$r$(.)]$p$(wind+1st)	&8 &0	\\
B.2. NB[$\Lambda$(.)$\alpha$(.)]Exponential[$r$(.)]$p$(wind) &7 &0.9\\
B.3. NB[$\Lambda$(.)$\alpha$(.)]Exponential[$r$(.)]$p$(1st)	&5	&5.0
\\B.4. NB[$\Lambda$(.)$\alpha$(.)]Exponential[$r$(.)]$p$(.)	&4	&6.4\\
\hline
C. Dynamics && \\
C.1. NB[$\Lambda$(.)$\alpha$(.)]Ricker+Immigration[$r$(.)$K$(.)$\iota$(.)]$p$(wind+1st) &10	&0	\\
C.2. NB[$\Lambda$(.)$\alpha$(.)]Gompertz+Immigration[$r$(.)$K$(.)$\iota$(.)]$p$(wind+1st) &10	&8.4 \\
C.3. NB[$\Lambda$(.)$\alpha$(.)]Exponential+Immigration[$r$(.)$\iota$(.)]$p$(wind+1st) &9	&36.5\\
C.4. NB[$\Lambda$(.)$\alpha$(.)]Geometric-recruitment+Immigration[$\gamma$(.)$\omega$(.)$\iota$(.)]$p$(wind+1st) &10	&38.6\\
C.5. NB[$\Lambda$(.)$\alpha$(.)]Gompertz[$r$(.)$K$(.)]$p$(wind+1st) &9	&192.8\\
C.6. NB[$\Lambda$(.)$\alpha$(.)]Ricker[$r$(.)$K$(.)]$p$(wind+1st) &9	&195.1\\
C.7. NB[$\Lambda$(.)$\alpha$(.)]Exponential[$r$(.)]$p$(wind+1st)	&8 &271.3\\
C.8. NB[$\Lambda$(.)$\alpha$(.)]Geometric-recruitment[$\gamma$(.)$\omega$(.)]$p$(wind+1st) &9	&273.7\\
C.9. NB[$\Lambda$(.)$\alpha$(.)]Constant-recruitment[$\gamma$(.)$\omega$(.)]$p$(wind+1st) &9	&1856.7\\
\hline
\end{tabular}
\end{table}
\clearpage

%[Andy and Richard, which do you prefer for parameter estimates?  
%Table or figure?  Either one may need a bit of work still.]
%% latex table generated in R 3.0.0 by xtable 1.7-1 package
%% Wed Sep 18 14:26:41 2013
%\begin{table}[ht]
%\centering
%  \caption{Parameter estimates for ovenbirds in Maryland and Virginia from BBS data, 1966-2010.  
%MLE estimates come from the NB[$\Lambda$(.)$\alpha$(.)]Ricker+Immigration[$r$(.)$K$(.)$\iota$(.)]$p$(wind+1st) 
%model run in the \textbf{R} package \texttt{unmarked}; base estimates from the same model run in \textbf{JAGS};
%obs estimates from the base model with random observer effects added; and ES estimates from the base
%model with regional environmental stochasticity added.  Detectability parameters (intercept: $p_{0}$, 
%effect of wind speed 1: $p_{1}$, effect of wind speed 2: $p_{2}$, effect of wind speed 3 or higher: $p_{3}$, 
%effect of first run of a route by an observer: $p_{1st}$, and random observer effect: $\sigma_{p}$) are on the 
%logit scale.}
%\begin{tabular}{rllllll}
%  \hline
% & Parameter & MLE & base & obs & ES & ES and obs \\ 
%  \hline
%1 & $\Lambda$ & 33.4 $\pm$ 4.7 & 31.1 $\pm$ 5.2 & 42.5 $\pm$ 7.3 & 33.2 $\pm$ 5.8 & 43.8 $\pm$ 7.9 \\ 
%  2 & $\alpha$ & 0.654 $\pm$ 0.087 & 0.624 $\pm$ 0.108 & 0.730 $\pm$ 0.119 & 0.649 $\pm$ 0.110 & 0.753 $\pm$ 0.124 \\ 
%  3 & $r$ & 0.026 $\pm$ 0.006 & 0.011 $\pm$ 0.005 & 0.018 $\pm$ 0.006 & 0.020 $\pm$ 0.010 & 0.023 $\pm$ 0.010 \\ 
%  4 & $K$ & 56.9 $\pm$ 9.4 & 55.8 $\pm$ 18.2 & 101.8 $\pm$ 21.1 & 112.0 $\pm$ 58.8 & 141.5 $\pm$ 69.5 \\ 
%  5 & $\iota$ & 0.283 $\pm$ 0.043 & 0.333 $\pm$ 0.043 & 0.306 $\pm$ 0.051 & 0.329 $\pm$ 0.046 & 0.303 $\pm$ 0.052 \\ 
%  6 & $p_{0}$ & -1.492 $\pm$ 0.100 & -1.476 $\pm$ 0.097 & -1.994 $\pm$ 0.116 & -1.571 $\pm$ 0.099 & -2.056 $\pm$ 0.124 \\ 
%  7 & $p_{1}$ & -0.012 $\pm$ 0.021 & -0.011 $\pm$ 0.021 & -0.014 $\pm$ 0.021 & -0.013 $\pm$ 0.021 & -0.015 $\pm$ 0.021 \\ 
%  8 & $p_{2}$ & -0.075 $\pm$ 0.027 & -0.071 $\pm$ 0.027 & -0.056 $\pm$ 0.026 & -0.069 $\pm$ 0.027 & -0.055 $\pm$ 0.026 \\ 
%  9 & $p_{3}$ & -0.133 $\pm$ 0.049 & -0.135 $\pm$ 0.049 & -0.145 $\pm$ 0.048 & -0.141 $\pm$ 0.049 & -0.144 $\pm$ 0.048 \\ 
%  10 & $p_{1st}$ & -0.058 $\pm$ 0.028 & -0.047 $\pm$ 0.027 & -0.075 $\pm$ 0.029 & -0.045 $\pm$ 0.028 & -0.075 $\pm$ 0.029 \\ 
%  31 & $\sigma_{p}$ & - & - & 0.353 $\pm$ 0.034 & - & 0.349 $\pm$ 0.034 \\ 
%  42 & $\sigma_{\nu}$ & - & - & - & 0.074 $\pm$ 0.014 & 0.064 $\pm$ 0.014 \\ 
%   \hline
%\end{tabular}
%\end{table}

\clearpage
\begin{figure}
  \centering
  \includegraphics[width=6.5in]{../figs/oven_par_est}
\caption{Parameter estimates for ovenbirds in Maryland and Virginia from BBS data, 1966-2010.  
MLE estimates come from the NB[$\Lambda$(.)$\alpha$(.)]Ricker+Immigration[$r$(.)$K$(.)$\iota$(.)]$p$(wind+1st) 
model run in the \textbf{R} package \texttt{unmarked}; base estimates from the same model run in \textbf{JAGS};
obs estimates from the base model with random observer effects added; and ES estimates from the base
model with regional environmental stochasticity added.  Detectability parameters (intercept: $p_{0}$, 
effect of wind speed 1: $p_{1}$, effect of wind speed 2: $p_{2}$, effect of wind speed 3 or higher: $p_{3}$, 
effect of first run of a route by an observer: $p_{1st}$, and random observer SD: $\sigma_{p}$) are on the 
logit scale. Error bars show SE for MLE estimates and SD for all others.}
\label{fig:oven_par_est}
\end{figure}
